\documentclass{article}
\pagestyle{empty}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{multicol}
\setlength{\parindent}{0in}
\setlength{\parskip}{1em}
\begin{document}

\centerline{\large\bf CSCI 321, IF Game Specs, Spring 2016}
\centerline{\bf Due date: Wednesday, June 1, Midnight}

\begin{itemize}
  \item
Reread the 2D game specifications for general things I'll look for in
your game.  However, here I'm looking for a \textit{story}.  This, of
course, can mean almost anything, but in general the protagonist
(usually the player) overcomes obstacles in the course of some quest,
and in the end comes some surprising realization about herself or
himself, and either succeeds magnificently, or fails spectacularly. 

\item The source for your game should be more than 2000 words.  If you
  do a straightforward  spooky house type of story, about eight
  rooms with about four important objects in each room should be about
  right.  That's about 250 words per room, which should be enough to
  describe it and set up some puzzles.

\item There is no need for a user's manual for IF.  If there are
  special verbs you use in your game (like  ``spin the top'' or
  ``juggle the mice'') you should provide in-game hints that suggest
  these actions.  The presence of a top should automatically suggest
  spinning it.  Juggling could be suggested more subtly, perhaps with
  various juggler's paraphernalia about the room (colored balls and
  clubs, a unicycle, \textit{etc.}).  They can be locked in a glass
  cabinet if you don't want to provide interesting actions for all of
  them.  But then, why wouldn't you?

\item Clear out the skein until there are only the ``best'' paths
  through your game remaining.  Normally this will be only one path,
  but you could have two or more ``correct'' endings.  Make sure all
  are in the skein, and nothing else.

\item Bless the transcript of your best walkthrough.  If there is
  more than one correct ending, pick the best one.

\end{itemize}

  

\end{document}
\end
