\documentclass{beamer}
%\documentclass[handout]{beamer}

\usepackage{rotating}

\title{How to Be Creative}
\subtitle{Some Ideas from\\ {\em The Art of Game Design, a
  Book of Lenses}\\ Jesse Schell}
%\author{}
%\author[R. Rostamian]{Rouben Rostamian}
%\institute[WWU]{
%  Department of Computer Science \\
%  Western Washington University \\
%  Bellingham, WA 98225
%}

\begin{document}

%----------- titlepage ----------------------------------------------%
\begin{frame}[plain]
  \titlepage
\end{frame}

%----------- slide --------------------------------------------------%
\begin{frame}
  \frametitle{The Creative Process}


\begin{enumerate}
\item Think of an idea.
\item Try it out.
\item Keep changing it and testing it until it seems good enough.
\end{enumerate}

\end{frame}
%----------- slide --------------------------------------------------%
\begin{frame}
  \frametitle{Inspiration comes from experience}

\begin{itemize}
\item Stop looking at games and look {\em everywhere else.}\pause
\begin{itemize}
\item Playing other video games will make your game look the same.\pause
\item Watching TV will make your games look like TV shows.\pause
\end{itemize}
\item If you like to go to bars, go to church.\pause
\item If you're not athletic, play a physical game.\pause
\item If you are athletic, read a book.\pause
\item If you like classical, listen to pop.\pause
\item What experience in my life would I want to share with
  others?\pause
\item In what small way can I capture the essence of that experience
  in a game?\pause
\item What parts of the game could I use?
\begin{itemize}\item technology\item mechanics\item story\item aesthetics
\end{itemize}

\end{itemize}

\end{frame}

%----------- slide --------------------------------------------------%
\begin{frame}
  \frametitle{Focus on Stating the Problem}

\begin{itemize}
\item Examples:
\begin{itemize}
\item How can I make a board game that uses magnets?\pause
\item How can I make a videogame about Hansel and Gretel?\pause
\item How can I make a game that feels like a surrealist painting?\pause
\item How can I improve on Tetris?\pause
\end{itemize}
\item Most people jump to solutions too fast.\pause
\item What problems are you trying to solve?\pause
\item Have you been making assumptions about the game that have
  nothing to do with its true purpose?\pause
\item Is the game you're imagining really the best solution?\pause
\item How can you tell when you've solved the problem?
\end{itemize}

\end{frame}

%----------- slide --------------------------------------------------%
\begin{frame}
  \frametitle{Connect with Sleep}

\begin{itemize}
\item Salvador Dali:  slumber with a key.\pause
\item Write down your ideas when you first wake up.
\end{itemize}

\end{frame}


%----------- slide --------------------------------------------------%
\begin{frame}
  \frametitle{Your Creative Self: A silent partner}

\begin{itemize}
\item Can't talk\pause
\item Impulsive\pause
\item Emotional\pause
\item Playful\pause
\item Irrational

\end{itemize}

\end{frame}


%----------- slide --------------------------------------------------%
\begin{frame}
  \frametitle{Getting in Touch with your Subconscious}

\begin{itemize}
\item Pay attention\pause
\item Record your ideas\pause
\item Manage its appetites
\begin{itemize}
\item take care of food, safety, personal relationships
\end{itemize}
\pause
\item Sleep\pause
\item Don't push too hard
\begin{itemize}
\item ask a question and leave it alone
\end{itemize}
\end{itemize}

\end{frame}


%----------- slide --------------------------------------------------%
\begin{frame}
  \frametitle{Maslow's Hierarchy of Needs}

\begin{itemize}

\item Physiological
\begin{itemize}
\item Food, water, shelter, warmth
\end{itemize}

\item Safety
\begin{itemize}
\item Security, stability, freedom from fear
\end{itemize}

\item Belonging-Love
\begin{itemize}
\item Friends, family, spouse, lover
\end{itemize}

\item Self-esteem
\begin{itemize}
\item Achievement, mastery, recognition, respect
\end{itemize}

\item Self-actualization
\begin{itemize}
\item Creativity, fulfillment, inner talent
\end{itemize}

\end{itemize}

\end{frame}



%----------- slide --------------------------------------------------%
\begin{frame}
  \frametitle{Brainstorming Tips}

\begin{columns}
\column{0.5\textwidth}
\begin{itemize}
\item Write them down\pause
\item Write instead of type\pause
\item Sketch\pause
\item Toys\pause
\item Change your perspective\pause
\item Immerse yourself\pause
\item Crack jokes\pause
\item Spare no expense\pause
\end{itemize}
\column{0.5\textwidth}
\begin{itemize}
\item Write on the wall\pause
\item The space remembers\pause
\item Write everything\pause
\item Number your lists\pause
\item Mix and match categories\pause
\item Talk to yourself\pause
\item Find a partner\pause
\end{itemize}
\end{columns}
\end{frame}



%----------- slide --------------------------------------------------%
\begin{frame}
  \frametitle{Mix and Match}

\begin{columns}[t]
\column{.25\textwidth}
Technology
\begin{enumerate}\scriptsize
\item Mouse and keyboard
\item Game controller
\item Gesture
\item Cel phone
\item Handheld game
\item Use with Instant Messaging
\item Game console
\end{enumerate}
\column{.25\textwidth}
Mechanics
\begin{enumerate}\scriptsize
\item Sims-like game
\item Interactive fiction
\item The winner makes the most friends
\item Try to spread rumors about the other players
\item Try to help as many people as possible
\item Tetris-like game
\end{enumerate}
\column{.25\textwidth}
Story
\begin{enumerate}\scriptsize
\item High school drama
\item College-themed
\item You play cupid
\item You're a TV star
\item Hospital theme
\item You're a rock star
\item You're a dancer
\end{enumerate}
\column{.25\textwidth}
Aesthetic
\begin{enumerate}\scriptsize
\item Cell shading
\item Anime style
\item All characters are animals
\item R\&B music
\item Punk/rock music
\item Steampunk
\end{enumerate}
\end{columns}
\pause
\hfill {\em And make up new categories!}

\end{frame}




\end{document}
